### 2009-11-18 doc/tech_doc/sec/arc-security-documentation.tex
More references fixed. 2 teporary pictures added. Few mising pictures identified.

The ARC NOX defines specific evaluation request and policy schema. Based on the schema definition, one policy evaluation engine is implemented. The design principal of policy evaluation engine is generality by which the implementation of the policy evaluation engine can be easily extended to adopt some other policy schema, such as XACML policy schema.

@@

\subsection{Policy evaluation engine --- Support of XACML policy and request} % (fold)
\label{subsec:policy_engine_xacml}
Currently, XACML specification is partially supported/implemented in ARC. More specifically, except the \textit{$<$Obligation$>$} element, other elements are supported.

@@

\section{Policy Decision Service (Charon Service)} % (fold)
\label{sec:policy_decision_service}
Policy decision service is a service implementation which configurably contains the functionality of ArcPDP or XACMLPDP. It will accept the soap request containing policy decision request and return soap response containing policy decision response.

@@

It's configuration is presented in section \ref{subsec:pdpservice_conf}.

@@

\section{Security Attributes -- how to compose policy decision request to policy evaluation engine} % (fold)
\label{sec:sec_attributes}

@@

Each Security Attribute stores it's information in internal format and is capable to export it to one of predefined formats using \textit{Export()} method. Currently only supported format is ARC Policy/Request XML document described in sections \ref{subsec:authz_policy} and \ref{subsec:authz_request}.
Collectors of Security Attributes instantiate corresponding classes and link them to Secuirity Attributes containers - MessageAuth $<$arc/message/MessageAuth.h$>$ and MessageAuthContext $<$arc/message/Message.h$>$ storing collected attributes per request and per session correspondingly. Each attribute is assigned a name. Current implementations of Security Attributes Collectors are either integrated into existing MCCs or implemented as separate SecHandler plugins. See section \ref{subsec:sec_attr_avail_collectors} for available Collectors and corresponding Security Attributes.

@@

\subsubsection{TCP} % (fold)
\label{subsubsec:sec_attr_TCP}
Information is collected inside TCP MCC. The Security Attribute is stored under name 'TCP' and exports ARC Request with following attributes:

\begin{table}[ht]
\caption{Security Attributes collected at TCP MCC}
\centering
\begin{tabular}{| l | p{7cm} | p{5cm} |}
\hline
\textbf{Element} & \textbf{AttributeId} & \textbf{Content} \\ \hline
Resource & http://www.nordugrid.org/schemas/policy-arc/types/localendpoint & service\_ip[:service\_port] \\ \hline
SubjectAttribute & http://www.nordugrid.org/schemas/policy-arc/types/remoteendpoint & client\_ip[:client\_port] \\ \hline
\end{tabular}
\label{table:tcp_attr}
\end{table}

\subsubsection{TLS and VOMS} % (fold)
\label{subsubsec:sec_attr_TLS_VOMS}
Information is collected inside TLS MCC. Generated Security Attribute class is stored under name 'TLS' and exports ARC Request with following attributes:

\begin{table}[ht]
\caption{Security Attributes collected at TLS MCC}
\centering
\begin{tabular}{| l | p{7cm} | p{5cm} |}
\hline
\textbf{Element} & \textbf{AttributeId} & \textbf{Content} \\ \hline
SubjectAttribute & http://www.nordugrid.org/schemas/policy-arc/types/tls/ca & signer of first certificate in client's chain
 \\ \hline
SubjectAttribute & http://www.nordugrid.org/schemas/policy-arc/types/tls/chain & Subject of certificate in client's chain - multiple items \\ \hline
SubjectAttribute & http://www.nordugrid.org/schemas/policy-arc/types/tls/subject & Subject of last certificate in client's chain \\ \hline
SubjectAttribute & http://www.nordugrid.org/schemas/policy-arc/types/tls/identity & Subject of last non-proxy certificate in client's chain \\ \hline
SubjectAttribute & http://www.nordugrid.org/schemas/policy-arc/types/tls/vomsattribute & VOMS attributes extracted from whole client's chain of certificates \\ \hline
\end{tabular}
\label{table:tls_attr}
\end{table}

@@

If VOMS extensions contain user definable attributes those are presented together with the information of their grantor. They consist of \textit{voname} and \textit{hostname} elements presented above (if \textit{hostname} is missing it is assigned string value NULL) followed by user attribute. It's pattern is \textit{qualifier:name=value}. Here qualifier acts as namespace and is usually same as VO name. Below is an example of such attribute:

@@

\subsubsection{HTTP} % (fold)
\label{subsubsec:sec_attr_HTTP}
Information is collected inside HTTP MCC. The Security Attribute is stored under name 'HTTP' and exports ARC Request with following attributes:

\begin{table}[ht]
\caption{Security Attributes collected at HTTP MCC}
\centering
\begin{tabular}{| l | p{7cm} | p{5cm} |}
\hline
\textbf{Element} & \textbf{AttributeId} & \textbf{Content} \\ \hline
Resource & http://www.nordugrid.org/schemas/policy-arc/types/http/path & HTTP path without host and port part \\ \hline
Action & http://www.nordugrid.org/schemas/policy-arc/types/http/method & HTTP method \\ \hline
\end{tabular}
\label{table:http_attr}
\end{table}

\subsubsection{SOAP} % (fold)
\label{subsubsec:sec_attr_SOAP}
Information is collected inside SOAP MCC. Security Attribute is stored under name 'SOAP' and exports ARC Request with following attributes:

\begin{table}[ht]
\caption{Security Attributes collected at SOAP MCC}
\centering
\begin{tabular}{| l | p{7cm} | p{5cm} |}
\hline
\textbf{Element} & \textbf{AttributeId} & \textbf{Content} \\ \hline
Resource & http://www.nordugrid.org/schemas/policy-arc/types/soap/endpoint & To element of WS-Addressing structure \\ \hline
Action & http://www.nordugrid.org/schemas/policy-arc/types/soap/operation & SOAP top level element name without namespace prefix \\ \hline
Context & http://www.nordugrid.org/schemas/policy-arc/types/soap/namespace & Namespace of SOAP top level element \\ \hline
\end{tabular}
\label{table:soap_attr}
\end{table}

@@

\subsection{Delegation Architecture} % (fold)
\label{subsec:delegation_arch}
In current implementation delegation is achieved through Identity Delegation implemented using X.509 Proxy Certificates as defined in RFC 3820 \cite{x509proxy}. Client wishing to allow service to act on it's behalf provides Proxy Certificate to the service using Web Service based Delegation interface described in section \ref{subsec:delegation_interface}.

@@

\subsection{Delegation PDP} % (fold)
\label{subsec:delegation_pdp}
The Delegation PDP is similar to the Arc PDP described above except that it takes it's Policy documents directly from Security Attributes. Differently from Arc PDP it is meant to be used for enforcing policies defined by client. Configuration of Delegation PDP is described in section \ref{subsec:deleg_pdp_conf}.

@@

Delegation interface in the ARC NOX is implemented using Web Service approach. Each ARC NOX service wishing to act on behalf of client identity implements this interface in order to accept delegated credentials. Here is how delegation procedure works (also shown in Figure \ref{fig:delegation_flow} and Figure \ref{fig:delegation_flow1}) :

@@

    \begin{itemize}
        \item validityStart, validityEnd and validityPeriod  specify when Proxy becomes valid, when it's validity finishes or for how long the Proxy is valid respectively. For example

@@

    In terms of implementation, since the SSO profile is primarily supposed to protect Web applications and provide authentication for Web users, some external code on the client and service side is implemented to integrate the SSO profile. On the client side, apart from the client interface for writing Web Service client, the user agent functionality of the Web browser is implemented in order to mimic its behavior, such as HTTP redirection and HTTP cookie processing. In fact, implementation of the user agent is also based on the client interface of ARC, specifically, the HTTPs client interface, since the client interface of ARC can support different protocols which are incarnated by different MCCs. The client developers who would use SAML2.0 SSO profile should call the user agent interface and then the Web Service client interface. On the service side, the Service Provider functionality (based on the HTTP MCC configured together with TLS MCC) is implemented by SP Service. For Identity Provider, the Shibboleth IdP implementation is used. Figure 8 shows the process of SAML2.0 SSO integrated in ARC client and service.

@@

The steps shows in Figure 8 are described as follows:

@@

    The message flow of interaction between Client, ARC SP Service, ARC Service and Shibboleth IdP is also shown in Figure.9.

@@

    The SLCS service is also a Web Service (standard Web Service implemented by using ARC service interface), and the SLCS client is a specific command-line interface (CLI) which includes the user agent and WS client. The whole process of SLCS invocation is showed in Figure 10 (from step 1 to step 8), which is the same as in Figure 8, except that step 7 and step 8 are invoked for the SLCS certificate request and response.

 SLCS service is supposed to run together with SP service. A typical configuration for SLCS service is as follows:

@@

    SLCS service and SP service should together be configured by using the Plexer. SP service is directly based on http and SLCS is directly based on soap. As shown in the above table, a plexer dispatches the message flow (outgoing from http.service) into two SP service (saml.sp) and SOAP service (soap.service).

@@

    Since the lifetime of the short lived credential is normally short, it is not a must to protect the private key by a pass phrase. As illustrated in steps (a) and (b) in Figure 10, if the private key is not protected through the Web Service client, the user can use the X.509 certificate to access Grid Service or Web Service from any kind of middleware. If the private key is protected, she can use the X.509 certificate to generate a proxy certificate (by using a command-line interface utility such as grid-proxy-init , voms-proxy-init, or arcproxy ), and then use the proxy certificate to access a Grid/Web Service.

@@

    As shown in Figure 11, there is a specific WS client and Web Service for processing delegation: delegation client, and delegation service. The WSDL (Web Service Description Language) of the delegation service has three main operations, the first one for processing the delegation initiation, the second one for storing the signed proxy certificate, and the last one for processing acquisition of the delegated credential. The sequential diagram of the delegation is detailed in Figure 12.

    For the delegation process, the delegation WS client functionality (inside each delegation service or the initial client) which is included to access the target delegation service (d1 or d2 in Figure 11), should use the user’s X.509 credential or a delegated credential (on behalf of the user) for authentication and secure communication, as well as delegate one more level of the delegated credential. For the other general  service invocation process, the WS client functionality (inside each service or the initial client) which is included to access the target service (step a1 or a2 in Figure 11), should also use the user’s X.509 credential or a delegated credential for authentication and secure communication.

    Since each WS client accesses the target service (delegation service or other general services) on behalf of the user, the trust relationship between the user’s certificates (X.509 certificate and proxy certificate) and services is required, while the trust relationship between service’s certificate and another service’s certificate is not required.

>>>>

The ARC NOX defines specific evaluation request and policy schema. Based on the schema definition the policy evaluation engine is implemented. The design principal of policy evaluation engine is generality by which the implementation of the policy evaluation engine can be easily extended to adopt some other policy schema, such as XACML policy schema.

@@

\subsection{Policy evaluation engine. Support for XACML policy and request} % (fold)
\label{subsec:policy_engine_xacml}
Currently, XACML specification \cite{xacml} is partially supported/implemented in ARC. More specifically, except the \textit{$<$Obligation$>$} element, other elements are supported.

@@

\section{Policy Decision Service (Charon Service)} % (fold)
\label{sec:policy_decision_service}
Policy decision service is a service implementatiing Arc PDP and XACML PDP depending on its configuration. It will accept the SOAP request containing policy decision request and return SOAP response containing policy decision response.

@@

Its configuration is presented in section \ref{subsec:pdpservice_conf}.

@@

\section{Security Attributes. How to compose policy decision request to policy evaluation engine} % (fold)
\label{sec:sec_attributes}

@@

Each Security Attribute stores its information in internal format and is capable to export it to one of predefined formats using \textit{Export()} method. Currently only supported format is ARC Policy/Request XML document described in sections \ref{subsec:authz_policy} and \ref{subsec:authz_request}.
Collectors of Security Attributes instantiate corresponding classes and link them to Secuirity Attributes containers - MessageAuth $<$arc/message/MessageAuth.h$>$ and MessageAuthContext $<$arc/message/Message.h$>$ storing collected attributes per request and per session correspondingly. Each attribute is assigned a name. Current implementations of Security Attributes Collectors are either integrated into existing MCCs or implemented as separate SecHandler plugins. See section \ref{subsec:sec_attr_avail_collectors} for available Collectors and corresponding Security Attributes.

@@

\subsubsection{TCP} % (fold)
\label{subsubsec:sec_attr_TCP}
Information is collected inside TCP MCC. The Security Attribute is stored under name 'TCP' and exports ARC Request with attributes described in table \ref{table:tcp_attr}.

\begin{table}[h]
\caption{Security Attributes collected at TCP MCC}
\centering
\begin{tabular}{| l | p{7cm} | p{5cm} |}
\hline
\textbf{Element} & \textbf{AttributeId} & \textbf{Content} \\ \hline
Resource & http://www.nordugrid.org/schemas/policy-arc/types/localendpoint & service\_ip[:service\_port] \\ \hline
SubjectAttribute & http://www.nordugrid.org/schemas/policy-arc/types/remoteendpoint & client\_ip[:client\_port] \\ \hline
\end{tabular}
\label{table:tcp_attr}
\end{table}

\subsubsection{TLS and VOMS} % (fold)
\label{subsubsec:sec_attr_TLS_VOMS}
Information is collected inside TLS MCC. Generated Security Attribute class is stored under name 'TLS' and exports ARC Request with attributes described in table \ref{table:tls_attr}.

\begin{table}[h]
\caption{Security Attributes collected at TLS MCC}
\centering
\begin{tabular}{| l | p{7cm} | p{5cm} |}
\hline
\textbf{Element} & \textbf{AttributeId} & \textbf{Content} \\ \hline
SubjectAttribute & http://www.nordugrid.org/schemas/policy-arc/types/tls/ca & signer of first certificate in client's chain
 \\ \hline
SubjectAttribute & http://www.nordugrid.org/schemas/policy-arc/types/tls/chain & Subject of certificate in client's chain - multiple items \\ \hline
SubjectAttribute & http://www.nordugrid.org/schemas/policy-arc/types/tls/subject & Subject of last certificate in client's chain \\ \hline
SubjectAttribute & http://www.nordugrid.org/schemas/policy-arc/types/tls/identity & Subject of last non-proxy certificate in client's chain \\ \hline
SubjectAttribute & http://www.nordugrid.org/schemas/policy-arc/types/tls/vomsattribute & VOMS attributes extracted from whole client's chain of certificates \\ \hline
\end{tabular}
\label{table:tls_attr}
\end{table}

@@

If VOMS extensions contain user definable attributes those are presented together with the information of their grantor. They consist of \textit{voname} and \textit{hostname} elements presented above (if \textit{hostname} is missing it is assigned string value NULL) followed by user attribute. Its pattern is \textit{qualifier:name=value}. Here qualifier acts as namespace and is usually same as VO name. Below is an example of such attribute:

@@

\subsubsection{HTTP} % (fold)
\label{subsubsec:sec_attr_HTTP}
Information is collected inside HTTP MCC. The Security Attribute is stored under name 'HTTP' and exports ARC Request with attributes described in table \ref{table:http_attr}.

\begin{table}[h]
\caption{Security Attributes collected at HTTP MCC}
\centering
\begin{tabular}{| l | p{7cm} | p{5cm} |}
\hline
\textbf{Element} & \textbf{AttributeId} & \textbf{Content} \\ \hline
Resource & http://www.nordugrid.org/schemas/policy-arc/types/http/path & HTTP path without host and port part \\ \hline
Action & http://www.nordugrid.org/schemas/policy-arc/types/http/method & HTTP method \\ \hline
\end{tabular}
\label{table:http_attr}
\end{table}

\subsubsection{SOAP} % (fold)
\label{subsubsec:sec_attr_SOAP}
Information is collected inside SOAP MCC. Security Attribute is stored under name 'SOAP' and exports ARC Request with attributes described in table \ref{table:soap_attr}.

\begin{table}[h]
\caption{Security Attributes collected at SOAP MCC}
\centering
\begin{tabular}{| l | p{7cm} | p{5cm} |}
\hline
\textbf{Element} & \textbf{AttributeId} & \textbf{Content} \\ \hline
Resource & http://www.nordugrid.org/schemas/policy-arc/types/soap/endpoint & To element of WS-Addressing structure \\ \hline
Action & http://www.nordugrid.org/schemas/policy-arc/types/soap/operation & SOAP top level element name without namespace prefix \\ \hline
Context & http://www.nordugrid.org/schemas/policy-arc/types/soap/namespace & Namespace of SOAP top level element \\ \hline
\end{tabular}
\label{table:soap_attr}
\end{table}

@@

\subsection{Delegation Architecture} % (fold)
\label{subsec:delegation_arch}
In current implementation delegation is achieved through Identity Delegation implemented using X.509 Proxy Certificates as defined in RFC 3820 \cite{x509proxy}. Client wishing to allow service to act on its behalf provides Proxy Certificate to the service using Web Service based Delegation interface described in section \ref{subsec:delegation_interface}.

@@

\subsection{Delegation PDP} % (fold)
\label{subsec:delegation_pdp}
The Delegation PDP is similar to the Arc PDP described above except that it takes its Policy documents directly from Security Attributes. Differently from Arc PDP it is meant to be used for enforcing policies defined by client. Configuration of Delegation PDP is described in section \ref{subsec:deleg_pdp_conf}.

@@

Delegation interface in the ARC NOX is implemented using Web Service approach. Each ARC NOX service wishing to act on behalf of client identity implements this interface in order to accept delegated credentials. Here is how delegation procedure works (also shown in figure \ref{fig:delegation_flow} and figure \ref{fig:delegation_flow1}) :

@@

    \begin{itemize}
        \item validityStart, validityEnd and validityPeriod  specify when Proxy becomes valid, when its validity finishes or for how long the Proxy is valid respectively. For example

@@

    In terms of implementation, since the SSO profile is primarily supposed to protect Web applications and provide authentication for Web users, some external code on the client and service side is implemented to integrate the SSO profile. On the client side, apart from the client interface for writing Web Service client, the user agent functionality of the Web browser is implemented in order to mimic its behavior, such as HTTP redirection and HTTP cookie processing. In fact, implementation of the user agent is also based on the client interface of ARC, specifically, the HTTPs client interface, since the client interface of ARC can support different protocols which are incarnated by different MCCs. The client developers who would use SAML2.0 SSO profile should call the user agent interface and then the Web Service client interface. On the service side, the Service Provider functionality (based on the HTTP MCC configured together with TLS MCC) is implemented by SP Service. For Identity Provider, the Shibboleth IdP implementation is used. Figure \ref{fig:saml_sso} shows the process of SAML2.0 SSO integrated in ARC client and service.

@@

\begin{figure}[ht]
\centering{{{\includegraphics[width=0.5\textwidth]{saml_sso.png}}}
\caption{\label{fig:saml_sso}SAML 2.0 SSO Profile in ARC} }
\end{figure}

The steps shows in figure \ref{fig:saml_sso} are described as follows:

@@

    The message flow of interaction between Client, ARC SP Service, ARC Service and Shibboleth IdP is also shown in figure \ref{fig:saml_message_flow}.

\begin{figure}[ht]
 \centering{
  % {{\includegraphics[width=0.5\textwidth]{missing_picture.png}}}
  \caption{\label{fig:saml_message_flow}SAML 2.0 SSO Profile message flow}
}
\end{figure}

@@

    The SLCS service is also a Web Service (standard Web Service implemented by using ARC service interface), and the SLCS client is a specific command-line interface (CLI) which includes the user agent and WS client. The whole process of SLCS invocation is showed in figure \ref{fig:slcs} (from step 1 to step 8), which is the same as in figure \ref{fig:saml_sso}, except that step 7 and step 8 are invoked for the SLCS certificate request and response.

\begin{figure}[ht]
 \centering{
  {{\includegraphics[width=0.7\textwidth]{slcs.png}}}
  \caption{\label{fig:slcs}Short Lived Credential Service}
}
\end{figure}

 The SLCS service is supposed to run together with SP service. A typical configuration for SLCS service is as follows:

@@

    SLCS service and SP service should together be configured by using the Plexer. SP service is directly based on http and SLCS is directly based on SOAP. As shown in the above table, a plexer dispatches the message flow (outgoing from http.service) into two SP service (saml.sp) and SOAP service (soap.service).

@@

    Since the lifetime of the short lived credential is normally short, it is not a must to protect the private key by a pass phrase. As illustrated in steps (a) and (b) in figure \ref{fig:delegation_flow1}, if the private key is not protected through the Web Service client, the user can use the X.509 certificate to access Grid Service or Web Service from any kind of middleware. If the private key is protected, she can use the X.509 certificate to generate a proxy certificate (by using a command-line interface utility such as grid-proxy-init , voms-proxy-init, or arcproxy ), and then use the proxy certificate to access a Grid/Web Service.

@@

    As shown in figure \ref{fig:delegation_service}, there is a specific WS client and Web Service for processing delegation: delegation client, and delegation service. The WSDL (Web Service Description Language) of the delegation service has three main operations, the first one for processing the delegation initiation, the second one for storing the signed proxy certificate, and the last one for processing acquisition of the delegated credential. The sequential diagram of the delegation is detailed in figure \ref{fig:delegation_service_flow}.

\begin{figure}[ht]
 \centering{
  % {{\includegraphics[width=0.7\textwidth]{missing_picture.png}}}
  \caption{\label{fig:delegation_service}X.509 Credential Delegation Service}
}
\end{figure}

\begin{figure}[ht]
 \centering{
  % {{\includegraphics[width=0.7\textwidth]{missing_picture.png}}}
  \caption{\label{fig:delegation_service_flow}X.509 Credential Delegation message flow}
}
\end{figure}

    For the delegation process, the delegation WS client functionality (inside each delegation service or the initial client) which is included to access the target delegation service (d1 or d2 in figure \ref{fig:delegation_service}), should use the user’s X.509 credential or a delegated credential (on behalf of the user) for authentication and secure communication, as well as delegate one more level of the delegated credential. For the other general  service invocation process, the WS client functionality (inside each service or the initial client) which is included to access the target service (step a1 or a2 in figure \ref{fig:delegation_service}), should also use the user's X.509 credential or a delegated credential for authentication and secure communication.

    Since each WS client accesses the target service (delegation service or other general services) on behalf of the user, the trust relationship between the user's certificates (X.509 certificate and proxy certificate) and services is required, while the trust relationship between service's certificate and another service's certificate is not required.
